---
title       : CD breakfast 
subtitle    : With R
author      : Levente Otti
job         : R Hacker
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
github      :
  user: iamotti
  repo: rwed
---

## What is R?

> 1. Functional programming language
> 2. Statistical and data mining tool
> 3. Data Visualisation tool
> 4. Presentation tool (Knitr, Slides, Pdf, Html, ...)
> 5. Much more (munging, transformation, sampling, ...)

--- .class #id 

## Pros
> 1. Free, open source, avaliable on major platforms
> 2. Massive set of packages for statistical modeling, machine learning, ...
> 3. Fantastic community (mailing lists, stackoverflow, BURN, ...)
> 4. Powerful tools for communicating the results, html, reports, interactive websites...
> 5. Strong foundation in functioanal proramming
> 6. Powerful IDE for interactive data analysis and statistical programming
> 7. Designed to connect to high-performance programming languages like C++ 
> 8. "Everything" integrated and easy to use => quick results

---

## Cons

> 1. Most R users are not programmers (code is not very elegant, fast or easy to understand)
> 2. Focus on result instead of process (lack of VCS, TDD)
> 3. Inconsistency is rife across contributed packages, even within base R (20 years)
> 4. Many special cases
> 5. R is not fast, easy to write terribly slow code with profligate memory usage
> 6. All the above is a challenge for us

---

## Example: Simple Correlation 

```{r results='asis', echo=TRUE, message=FALSE, tidy=FALSE}
library(ggplot2)
qplot(speed, dist, data = cars) + geom_smooth()
```

---

## Example: Con - read.table
Readability, speed, hidden features (first column), ...
```{r results='asis', echo=TRUE, message=FALSE, tidy=FALSE}
read.table <- function (file, header = FALSE, sep = "", quote = "\"'", dec = ".", 
  row.names, col.names, as.is = !stringsAsFactors, na.strings = "NA", 
  colClasses = NA, nrows = -1, skip = 0, check.names = TRUE, 
  fill = !blank.lines.skip, strip.white = FALSE, blank.lines.skip = TRUE, 
  comment.char = "#", allowEscapes = FALSE, flush = FALSE, 
  stringsAsFactors = default.stringsAsFactors(), fileEncoding = "", 
  encoding = "unknown", text, skipNul = FALSE) 
{
  if (missing(file) && !missing(text)) {
    file <- textConnection(text, encoding = "UTF-8")
    encoding <- "UTF-8"
    on.exit(close(file))
  }
  if (is.character(file)) {
    file <- if (nzchar(fileEncoding)) 
      file(file, "rt", encoding = fileEncoding)
    else file(file, "rt")
    on.exit(close(file))
  }}
```

---

## Example: GoogleVis a more complex example
```{r results='asis', echo=FALSE, message=FALSE, tidy=FALSE}
library(googleVis)
G <- gvisGeoChart(Exports, "Country", "Profit", 
                  options=list(width=250, height=120), chartid="c1")
 
T <- gvisBarChart(Exports[,1:2], yvar="Profit", xvar="Country",                  
                  options=list(width=250, height=260, 
                               legend='none'), chartid="c2")
 
GT <- gvisMerge(G,T, horizontal=FALSE, chartid="gt") 
 
## Tree map
 
M <- gvisTreeMap(Regions,  "Region", "Parent", "Val", "Fac",
                    options=list(width=400, height=380,
                                 fontSize=16,
                                 minColor='#EDF8FB',
                                 midColor='#66C2A4',
                                 maxColor='#006D2C',
                                 headerHeight=20,
                                 fontColor='black',
                                 showScale=TRUE), chartid="c3")
 
GTM <- gvisMerge(GT, M, horizontal=TRUE,
                     tableOptions="cellspacing=5", chartid="gtm")
 
print(GTM, 'chart')
```

---

## Example: Song Kata - Test
```{r results='asis', echo=TRUE, message=FALSE, tidy=FALSE}
library(testthat)
context ("Complex example")

input <- "
    15 3
    197812 re_hash
    78906 5_4
    189518 tomorrow_comes_today 
  "

test_that("output contains the list of zips songs for 14 rows", {
  output <- as.data.frame(read.table(header = FALSE, col.names=c("name"), text = "
    19_2000
    clint_eastwood
    tomorrow_comes_today
  "))
  expect_that(get_zips_songs_list(input), equals(output))
})
```

---

## Example: Song Kata - Source
```{r results='asis', echo=TRUE, message=FALSE, tidy=FALSE}
get_zips_songs_list <- function (input) {
  result_count <- get_number_of_result_list(input)
  data <- read_input_into_dataframe(input)
  
  data$i <- as.numeric(rownames(data))
  data$qi <- data$fi/(1/data$i)
  head(arrange(data, desc(qi), i) %.% select(name),result_count)
}


get_number_of_result_list <- function (input) {
  as.numeric(as.character(as.data.frame(read.table(header = FALSE, text = input))[1,2]))
}



read_input_into_dataframe <- function(input) {
  as.data.frame(read.table(header = FALSE, text = input, col.names=c("fi","name"), skip=2))
}
```

---

## Example: GoogleVis vs HDS 
> 1. Aggregated by hour (request time, partition, filter count)
http://127.0.0.1:25127/custom/googleVis/MotionChartID22f6573927b.html
> 2. Aggregated by day (filter count, time, line chart)
http://127.0.0.1:25127/custom/googleVis/MotionChartID22f695fd46.html

---

## Example: Interactivity
> 1. Interactive web site with Shiny
http://ottawa:5507 (http://192.168.8.24:5507)

---

## References
1. Hadley Wickham - Advanced R Programming (http://adv-r.had.co.nz/)
2. Ramnathv - Slidify (http://slidify.org/)
3. RStudio - Shiny (http://shiny.rstudio.com/) 

---

## Questions
and Answers

---